---
title: "Blackjack Hand Analysis"
output:
  html_notebook: default
  pdf_document: default
---

###### In this notebook, we attempt to recreate what is commonly known as "Basic Strategy" in blackjack.   
###### Basic strategy is a chart that recommends the best move (hit or stand) for each possible hand.  
###### The data used in this program can be derived from blackjack_hand_generator, which simulates a user-specified number of hands, and records various metrics about the hand, including the result, in csv files (stand.csv & hit.csv).    
  
  
```{r include=FALSE}
install.packages('dplyr')
library('dplyr')

install.packages("ggplot2")
library("ggplot2")
```



First, we must import the hand data, which is obtained using blackjack_hand_generator.py  
Here, we can see various metrics pertaining to each hand, including the players/dealers cards and totals, as well as the result of the hand:
```{r}
stand_hands = read.csv('stand.csv')
hit_hands = read.csv('hit.csv')

head(stand_hands)
head(hit_hands)
```

Next, we will filter the hands according to the result (W/L), so that we can determine the win percentage for each possible hand
```{r}
stands_won = filter(stand_hands, Result == 'W')
stands_lost = filter(stand_hands, Result == 'L')

hits_won = filter(hit_hands, Result == 'W')
hits_lost = filter(hit_hands, Result == 'L')

stands_won
stands_lost
hits_won
hits_lost
```


```{r}

```



Next, we create frequency tables for all hands played where the player stands, as well as a table for all of our standing wins.   
We can divide the 'win table' by the 'all_hands' table to derive a 'win percentage table'. 
```{r}
stand_hands = subset(stand_hands, Player_Total != 21)
stands_won = subset(stands_won, Player_Total != 21)
all_stand = table(stand_hands$Player_Total, stand_hands$DC1)
all_stand

stand_w = table(stands_won$Player_Total, stands_won$DC1)
stand_w 

stand_win_pct = round(((stand_w / all_stand) * 100), 2)
stand_win_pct
```

Next, we will repeat the same process with the hands where the player hits  
We can eliminate instances where the players total exceeds 21, because these hands automatically result in a loss.  
We will also eliminate instances of a player total of 2 so that we can have conformable matrices for our final pct table calculation. It is also inherently obvious that we should always hit in these instances because the dealer will always beat a player total of 2. 
```{r}

hit_hands = subset(hit_hands, player_open_total != 21)
#hit_hands[hit_hands$player_open_total !=2, ]

all_hit = table(hit_hands$player_open_total, hit_hands$DC1)

all_hit

hit_w = table(hits_won$player_open_total, hits_won$DC1)
hit_w 

hit_win_pct = round(((hit_w / all_hit) * 100), 2)
hit_win_pct

```

```{r}
hit_win_pct
stand_win_pct

final <- ifelse(hit_win_pct < stand_win_pct, yes = 'S', no = 'H')
final
```

```{r}
test = filter(stand_hands, DC1 == 1)
test
```

